19Q2

### 2019-07-19 19Q2.Rmd
update Rmd

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata, eval=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
ccl_lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])
lineage_converter <- unique(select(ccl_lineage_converter, grouped_general_doid, group_general_lineage_name))
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")

ccl_meta <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta <- select(ccl_meta, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
ccl_meta <- subset(ccl_meta, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " ", as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == "other" & CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid", disease)))))))

ccl_meta <- merge(ccl_meta, ccl_histologies, by = c("Cell_line", "CCLE_histology", "DepMap_ID"), all = TRUE)
ccl_meta$disease <- with(ccl_meta, ifelse(is.na(disease) & !is.na(CCLE_histology), CCLE_histology, disease))
ccl_meta$disease <- with(ccl_meta, ifelse(CCLE_histology == "large_intestine", "colorectal", 
                                   ifelse(CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid",
                                   ifelse(CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(CCLE_histology == "oesophagus", "esophagus", 
                                   ifelse(CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(Cell_line == "GT3TKB", "lung",
                                   ifelse(disease == "stomach" | CCLE_histology == "stomach", "gastric", disease))))))))

write.csv(ccl_meta, "./data_munging/19Q2/sample_info_edit_20190717.csv")
```

```{r}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20190717.csv", sep = ",", header = TRUE, row.names = 1)
ccl_meta <- ccl_meta %>% select(DepMap_ID, Cell_line, disease)
ccl_meta <- filter(ccl_meta, DepMap_ID != "ACH-002397")
```

@@

# Manually assign biomarker_types based on a combination of biomarker_type derived from the JSON string and patterns in the Feature.Names
biomarker_type_curated <- 
  ifelse(!is.na(str_match(string = tolower(g2p$biomarker_type), pattern = "copy number variant|amplification")[,1]) | !is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "copy number variant|amplification")), "Amplification/CNV",
  ifelse(!is.na(str_match(string = toupper(g2p$Feature.Names), pattern = "[A-Z0-9|A-Z]+-[A-Z0-9|A-Z]+|FUSION|POSITIVE")[,1]) & is.na(str_match(string = g2p$Feature.Names, pattern = "[0-9]+-[0-9]+")) | g2p$biomarker_type == "fusion" | g2p$biomarker_type == "rearrange", "Fusion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+([A-Z]|[*])|[0-9A-Z]+ [A-Z][0-9]+|[A-Z][0-9]+$")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "expression|methylation")[,1]), "Expression",
  ifelse(!is.na(str_match(string = g2p$biomarker_type, pattern = "intron|insertion|deletion|mutant")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "codon|positive|biallelic inactivation|del|ins|dup|mut|frameshift|missense|variant|[0-9]+fs|>")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "^KIT$|^BRCA$")[,1]), "Mutation", NA)))))))
g2p$biomarker_type_curated <- biomarker_type_curated[,1]

@@

# Add association columns
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p <- merge(lineage_converter, g2p, by = "grouped_general_doid", all.y = TRUE)
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p <- unique(g2p)
write.csv(g2p, "./data_munging/g2p_db_edited.csv")
```

```{r g2p}
g2p <- read.delim("./data_munging/g2p_db_edited.csv", sep = ",", header = TRUE, row.names = 1)
g2p$Drug_Gene_Lin <- gsub(pattern = " cancer", "", g2p$Drug_Gene_Lin)

@@

g2p$biomarker_type_curated <- ifelse(is.na(g2p$biomarker_type_curated), "None", as.character(g2p$biomarker_type_curated))
g2p_split <- filter(g2p, !is.na(Drug)) %>% group_split(biomarker_type_curated)
g2p_amp <- g2p_split[[1]]
g2p_expr <- g2p_split[[2]]
g2p_fus <- g2p_split[[3]]
g2p_fus$FusionPair <- sapply(X = g2p_fus$Feature.Names, function(x) strsplit(as.character(x)," ")[[1]][1])
g2p_fus$FusionPair <- ifelse(g2p_fus$FusionPair == "Fusions", as.character(g2p_fus$Gene), as.character(g2p_fus$FusionPair))
g2p_fus$FusionPair <- ifelse(g2p_fus$FusionPair == "BCR-ABL", "BCR-ABL1", as.character(g2p_fus$FusionPair))
g2p_fus$Drug_FusionPair <- ifelse(is.na(g2p_fus$Drug) | is.na(g2p_fus$FusionPair), NA, paste0(g2p_fus$Drug, "_", g2p_fus$FusionPair))
g2p_mut <- g2p_split[[4]] %>% drop_na(Direction_Curated)
g2p_mut_genes <- unique(sapply(strsplit(as.character(g2p_mut$Drug_Gene)[!is.na(g2p_mut$Drug_Gene)], "_", fixed = TRUE), function(x) x[2]))
g2p_nodrug <- filter(g2p, is.na(Drug))

g2p_prot <- unique(filter(g2p, !is.na(Gene_Protein_Change) & !is.na(Drug)) %>% select(group_general_lineage_name, Drug, Direction_Curated, Gene, Protein_Change, Gene_Protein_Change))
```

@@

```{r fusion_calls, eval=FALSE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
write.csv.gz(fusions, "./data_munging/19Q2/CCLE_fusions_edit_20190708.csv.gz")
```

```{r}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20190708.csv.gz", header = TRUE, sep = ",")
```

@@

```{r}
ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_dg_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)

@@

```{r wilcoxon_heatmap_function}
MakeWilcoxonHeatmap <- function(signif_df, feature, row_names, col_title, color, legend_title, legend_breaks, row_title_txt, radius, na_color) {
  if(missing(na_color)) {
    na_color <- "white"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "mut"))) {
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

@@

    signif_grid_annot <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot[signif_grid_annot == "resistance"] <- "X"
    signif_grid_annot[signif_grid_annot == "sensitivity"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) {
    signif_grid <- dcast(signif_df, FusionPair ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid < -log10(0.05)] <- NA

    signif_grid_annot <- dcast(signif_df, FusionPair ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  col_fun <- colorRamp2(c(legend_breaks[1], legend_breaks[length(legend_breaks)]), c("gray97", color))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70", at = legend_breaks, labels = legend_breaks)
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = na_color,
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  if(!is.na(str_match(string = tolower(feature), pattern = "prot"))) {
                    grid.text(signif_grid_annot1[i, j], x = x - width*0.20, y = y + height*0.07, gp = gpar(fontsize = 17))
                  }
                  if(signif_grid_annot[i, j] == "O") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "X") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "resistance") {
                    grid.circle(x = x - width*0.20, y = y + height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "both") {
                    grid.circle(x = x - width*0.20, y = y + height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                }
                )
  return(ht)
}
```

@@

### Lieage

@@

  col_fun <- colorRamp2(color_range, c("red3", "white", "blue3"))

>>>>

DepMap 19Q2 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r depmap_metadata, eval=FALSE}
ccl_converter <- read.delim("./data_munging/cell_line_database_v1_20180911.tsv", row.names = 1, sep = "\t", header = TRUE)
colnames(ccl_converter)[colnames(ccl_converter) == "Broad_ID"] <- "DepMap_ID"
lineage_converter <- unique(ccl_converter[, c("DepMap_ID", "grouped_general_doid", "group_general_lineage_name")])

# OLD CCL_META
ccl_histologies <- read.delim("./data_munging/ccl_histologies_metadata.csv", sep = ",", header = TRUE, row.names = 1)
colnames(ccl_histologies) <- c("Cell_line", "CCLE_histology", "TCGA_classification", "CCLE_name", "DepMap_ID")
ccl_meta <- read.delim("./data_munging/19Q2/sample_info.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_meta <- select(ccl_meta, DepMap_ID, stripped_cell_line_name, CCLE_name, disease)
colnames(ccl_meta) <- c("DepMap_ID", "Cell_line", "CCLE_name", "disease")
ccl_meta <- subset(ccl_meta, !grepl("MERGED", CCLE_name, fixed = TRUE))
ccl_diseases <- as.character(unique(ccl_meta$disease))[as.character(unique(ccl_meta$disease)) != " "]
CCLE_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_meta$CCLE_name, ignore.case = TRUE))
ccl_meta <- add_column(ccl_meta, CCLE_histology = CCLE_histology, .after = "CCLE_name")
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " ", as.character(CCLE_histology), as.character(disease)))
ccl_meta$disease <- with(ccl_meta, ifelse(disease == " " & CCLE_histology == "vulva", "vulva",
                                   ifelse(disease == " " & CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(disease == " " & CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(disease == " " & CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(disease == " " & CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(disease == "other" & CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid", disease)))))))

ccl_meta <- merge(merge(ccl_meta, ccl_histologies, by = c("Cell_line", "CCLE_histology", "CCLE_name", "DepMap_ID"), all = TRUE), lineage_converter, by = "DepMap_ID", all = TRUE)
ccl_meta$disease <- with(ccl_meta, ifelse(is.na(disease) & !is.na(CCLE_histology), CCLE_histology, disease))
ccl_meta$disease <- with(ccl_meta, ifelse(CCLE_histology == "large_intestine", "colorectal",
                                   ifelse(CCLE_histology == "haematopoietic_and_lymphoid_tissue", "haematopoietic_and_lymphoid",
                                   ifelse(CCLE_histology == "upper_aerodigestive_tract", "upper_aerodigestive",
                                   ifelse(CCLE_histology == "oesophagus", "esophagus",
                                   ifelse(CCLE_histology == "biliary_tract", "bile_duct",
                                   ifelse(Cell_line == "GT3TKB", "lung",
                                   ifelse(disease == "stomach" | CCLE_histology == "stomach", "gastric", disease))))))))
write.csv(ccl_meta, "./data_munging/19Q2/sample_info_edit_20190718.csv")
```

```{r}
ccl_meta <- read.delim("./data_munging/19Q2/sample_info_edit_20190717.csv", sep = ",", header = TRUE, row.names = 1)
ccl_meta <- ccl_meta %>% select(DepMap_ID, Cell_line, disease)
ccl_meta <- filter(ccl_meta, DepMap_ID != "ACH-002397")
ccl_unique <- data.frame(disease = unique(ccl_meta$disease))
```

@@

# Manually assign biomarker_types based on a combination of biomarker_type derived from the JSON string and patterns in the Feature.Names
biomarker_type_curated <- 
  ifelse(!is.na(str_match(string = tolower(g2p$biomarker_type), pattern = "copy number variant|amplification")[,1]) | !is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "copy number variant|amplification")), "Amplification/CNV",
  ifelse(!is.na(str_match(string = toupper(g2p$Feature.Names), pattern = "[A-Z0-9|A-Z]+-[A-Z0-9|A-Z]+|FUSION|ALK POSITIVE")[,1]) & is.na(str_match(string = g2p$Feature.Names, pattern = "[0-9]+-[0-9]+")) | g2p$biomarker_type == "fusion" | g2p$biomarker_type == "rearrange", "Fusion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+([A-Z]|[*])|[0-9A-Z]+ [A-Z][0-9]+|[A-Z][0-9]+$")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "expression|methylation|erbb2  positive")[,1]), "Expression",
  ifelse(!is.na(str_match(string = g2p$biomarker_type, pattern = "intron|insertion|deletion|mutant")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = tolower(g2p$Feature.Names), pattern = "codon|positive|biallelic inactivation|del|ins|dup|mut|frameshift|missense|variant|[0-9]+fs|>")[,1]), "Mutation",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "^KIT$|^BRCA$")[,1]), "Mutation", NA)))))))
g2p$biomarker_type_curated <- biomarker_type_curated[,1]

@@

# Curate lineage column
g2p <- separate_rows(g2p, Phenotype.Description, sep = ";")
g2p_lineages <- cbind(unique(select(g2p, Phenotype.Description, Phenotype.Family)), disease = NA)
g2p_lineages$disease <- with(g2p_lineages, ifelse(!is.na(str_match(string = tolower(Phenotype.Family), pattern = "astrocytoma|central nervous system|glioma")[,1]) | !is.na(str_match(string = tolower(Phenotype.Description), pattern = "astrocytoma|glioblastoma|craniopharyngioma|medulloblastoma")[,1]), "central_nervous_system",
                                           ifelse(Phenotype.Family %in% c("breast cancer") | Phenotype.Description %in% c("breast cancer", "Her2-receptor positive breast cancer", "invasive ductal carcinoma"), "breast",
                                           ifelse(!is.na(str_match(string = tolower(Phenotype.Family), pattern = "lymphoma|bone marrow|hematologic|immune system|leukemia")[,1]) | !is.na(str_match(string = tolower(Phenotype.Description), pattern = "leukemia|lymphoproliferative|myelodisplasic|systemic mastocytosis|hyper eosinophilic|waldenström macroglobulinemia|langerhans")[,1]), "haematopoietic_and_lymphoid",
                                           ifelse(Phenotype.Family %in% c("kidney cancer", "Nephrotic syndrome") | !is.na(str_match(string = tolower(Phenotype.Description), pattern = "renal|kidney")[,1]), "kidney",
                                           ifelse(Phenotype.Family %in% c("liver cancer"), "liver",
                                           ifelse(Phenotype.Family %in% c("lung cancer", "respiratory system cancer"), "lung",
                                           ifelse(Phenotype.Description %in% c("congenital fibrosarcoma", "synovial sarcoma", "myxoid liposarcoma", "liposarcoma"), "soft_tissue",
                                           ifelse(Phenotype.Family %in% c("integumentary system cancer", "skin cancer") | Phenotype.Description %in% c("basal cell carcinoma", "Basal cell carcinoma", "Dermatofibrosarcoma", "dermatofibrosarcoma protuberans", "melanoma", "skin melanoma", "squamous cell carcinoma"), "skin",
                                           ifelse(Phenotype.Family %in% c("peripheral nervous system neoplasm"), "peripheral_nervous_system",
                                           ifelse(Phenotype.Family %in% c("urinary bladder cancer"), "urinary_tract", 
                                           ifelse(!is.na(str_match(string = tolower(Phenotype.Description), pattern = "thyroid|papillary carcinoma")[,1]), "thyroid", 
                                           ifelse(Phenotype.Description %in% c("ovarian cancer"), "ovary", 
                                           ifelse(Phenotype.Description %in% c("rectal neoplasm"), "colorectal",
                                           ifelse(Phenotype.Description %in% c("oropharynx cancer"), "upper_aerodigestive", 
                                           ifelse(!is.na(str_match(string = tolower(Phenotype.Description), pattern = "gastrointestinal")[,1]), "gastric", NA))))))))))))))))

g2p <- merge(g2p, g2p_lineages, by = c("Phenotype.Description", "Phenotype.Family"), all = TRUE)

# Add association columns
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
# g2p <- merge(unique(select(ccl_meta, grouped_general_doid, disease)), g2p, by = "grouped_general_doid", all.y = TRUE)
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$disease), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$disease))
g2p <- unique(g2p)
write.csv(g2p, "./data_munging/g2p_db_edited_20190718.csv")
```

```{r g2p}
g2p <- read.delim("./data_munging/g2p_db_edited_20190718.csv", sep = ",", header = TRUE, row.names = 1)
g2p$Drug_Gene_Lin <- gsub(pattern = " cancer", "", g2p$Drug_Gene_Lin)

@@

g2p$biomarker_type_curated <- ifelse(is.na(g2p$biomarker_type_curated), "None", as.character(g2p$biomarker_type_curated))
g2p_split <- filter(g2p, !is.na(Drug)) %>% group_split(biomarker_type_curated)
g2p_amp <- g2p_split[[1]]
g2p_expr <- g2p_split[[2]]
g2p_fus <- g2p_split[[3]]
g2p_fus$FusionPair <- sapply(X = g2p_fus$Feature.Names, function(x) strsplit(as.character(x)," ")[[1]][1])
g2p_fus$FusionPair <- ifelse(g2p_fus$FusionPair == "Fusions", as.character(g2p_fus$Gene), as.character(g2p_fus$FusionPair))
g2p_fus$FusionPair <- ifelse(g2p_fus$FusionPair == "BCR-ABL", "BCR-ABL1", as.character(g2p_fus$FusionPair))
g2p_fus$Drug_FusionPair <- ifelse(is.na(g2p_fus$Drug) | is.na(g2p_fus$FusionPair), NA, paste0(g2p_fus$Drug, "_", g2p_fus$FusionPair))
g2p_mut <- g2p_split[[4]] %>% drop_na(Direction_Curated)
g2p_mut_genes <- unique(sapply(strsplit(as.character(g2p_mut$Drug_Gene)[!is.na(g2p_mut$Drug_Gene)], "_", fixed = TRUE), function(x) x[2]))

g2p_prot <- unique(filter(g2p, !is.na(Gene_Protein_Change) & !is.na(Drug)) %>% select(disease, Drug, Direction_Curated, Gene, Protein_Change, Gene_Protein_Change))
```

@@

```{r fusion_calls, eval=FALSE}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions.csv.gz", header = TRUE, sep = ",")
fusions_bcr_abl1 <- filter(fusions, X.FusionName %in% c("BCR--ABL1", "ABL1--BCR"))
fusions$LeftGene <- gsub(" .*", "", fusions$LeftGene)
fusions$RightGene <- gsub(" .*", "", fusions$RightGene)
fusions$X.FusionName <- gsub("--", "-", fusions$X.FusionName)
fusions <- select(fusions, DepMap_ID, X.FusionName, LeftGene, RightGene)
fusions <- as.data.frame(lapply(fusions, function(x) gsub("IGH@|IGH-@", "IGH", x)))
colnames(fusions) <- c("DepMap_ID", "FusionPair", "LeftGene", "RightGene")
fusions <- merge(fusions, ccl_meta, by = "DepMap_ID", all.x = TRUE)
write.csv.gz(fusions, "./data_munging/19Q2/CCLE_fusions_edit_20190718.csv.gz")
```

```{r}
fusions <- read.delim("./data_munging/19Q2/CCLE_fusions_edit_20190718.csv.gz", header = TRUE, sep = ",")
```

@@

```{r, eval=FALSE}
ccle_filt <- filter(ccle, Drug %in% g2p_drugs & Gene %in% g2p_dg_genes)
ccle_filt$Gene <- factor(ccle_filt$Gene, ordered = FALSE)
ccle_filt <- unfactorize(ccle_filt)

@@

```{r wilcoxon_heatmap_function}
MakeWilcoxonHeatmap <- function(signif_df, feature, row_names, col_title, color, legend_title, legend_breaks, row_title_txt, radius, na_color) {
  if(missing(na_color)) {
    na_color <- "white"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "mut"))) {
    signif_grid <- dcast(signif_df, Gene ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]

@@

    signif_grid_annot <- dcast(signif_df, Gene_Protein_Change ~ Drug, value.var = "Direction_Curated", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene_Protein_Change)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot[signif_grid_annot == "resistance"] <- "X"
    signif_grid_annot[signif_grid_annot == "sensitivity"] <- "O"
  }
  if(!is.na(str_match(string = tolower(feature), pattern = "fus"))) {
    signif_grid <- dcast(signif_df, FusionPair ~ Drug, value.var = "p_log10", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid <- signif_grid[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]

    signif_grid_annot <- dcast(signif_df, FusionPair ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-FusionPair)
    signif_grid_annot <- signif_grid_annot[, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
    signif_grid_annot[signif_grid_annot == "No"] <- ""
    signif_grid_annot[signif_grid_annot == "Yes"] <- "O"
  }
  col_fun <- colorRamp2(c(legend_breaks[1], legend_breaks[length(legend_breaks)]), c("gray97", color))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70", at = legend_breaks, labels = legend_breaks)
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = na_color,
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = row_title_txt,
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(signif_grid_annot1[i, j], x = x + width*0.25, y = y + height*0.07, gp = gpar(fontsize = 17))

                  if(signif_grid_annot[i, j] == "O") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "X") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "resistance") {
                    grid.circle(x = x - width*0.20, y = y + height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                  }
                  if(signif_grid_annot[i, j] == "both") {
                    grid.circle(x = x - width*0.20, y = y + height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = radius, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                  }
                }
                )
  return(ht)
}
```

@@

### Lineage

@@

  col_fun <- colorRamp2(color_range, c("forestgreen", "white", "blue3"))

  # Legend properties list
  ht_legend_list <- list(title_position = "topleft", legend_direction = "horizontal", legend_width = unit(3, "in"), border = "gray70", at = seq(color_range[1], color_range[3], by = 0.1), labels = seq(color_range[1], color_range[3], by = 0.1))
  ht <- Heatmap(corr_grid,
                col = col_fun,
                name = legend_title,
                na_col = "gray97",
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                row_title = "Genes",
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "top",
                cell_fun = function(j, i, x, y, width, height, fill) {
                  grid.text(signif_grid[i, j], x = x - width*0.20, y = y, gp = gpar(fontsize = 18))
                  if(annot_grid[i, j] == "sensitivity") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = height*0.15, default.units = "npc", gp = gpar(fill = "white", col = "black"))
                    # grid.circle(x = x, y = y, r = radius, gp = gpar(fill = "transparent", col = "black"))
                  }
                  if(annot_grid[i, j] == "resistance") {
                    grid.circle(x = x - width*0.20, y = y - height*0.20, r = height*0.15, default.units = "npc", gp = gpar(fill = "red", col = "black"))
                    # grid.text("☐", x, y, just = c(0.5, 0.425), gp = gpar(fontsize = 23, fontface = "bold"))
                  }
                }
                )
  return(ht)
}
```
